On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-02-05 main.tex
checked 2 and 3.1

% packages 
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{mathrsfs}
\usepackage{bm}
\usepackage{enumitem}
% new commands and shortcuts
\def\td{{\bf todo}}
\def\bs{\textit{\textbf{s}}}
\def\bz{\textit{\textbf{0}}}
\def\Lb{\textit{\textbf{L}}}
\def\Xb{\textit{\textbf{X}}}
\def\Lambdab{\bm{\Lambda}}
\def\Thetab{\bm{\Theta}}
\def\thetab{\bm{\vartheta}}
\def\mA{{\bm A}}
\def\fA{{\frak A}}
\def\xb{\textit{\textbf{x}}}
\def\fb{\textit{\textbf{f}}}
\def\ab{\textit{\textbf{a}}}
\def\pb{\textit{\textbf{p}}}
\def\ajb{\overline{a_j}}
\def\bjb{\overline{b_{j,l}}}
\def\pjb{\overline{P_{j}}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\htt}{ht}
\def\C{\mathbb{C}}
\def\pr{\mathbb{P}}
\def\vt{\vartheta}
\def\Q{\mathbb{Q}}
\def\R{\mathbb{R}}
\def\K{\mathbb{K}}
\def\rs{\mathscr{R}}
\def\P{\mathscr{P}}
\def\d{\delta}
\def\sing{ \textrm{sing}}
\def\codim{ \textrm{codim}}
\def\jac{ \textbf{jac}}
\def\exp{\textrm{exp}}
\def\grad{\textrm{grad}}
\def\rank{\textrm{rank}}
\def\reg{\textrm{reg}}
\def\jt{\widetilde{J}}
\def\At{\widetilde{A}}
\def\Yt{\widetilde{Y}}
\def\dtt{\widetilde{d}}
\def\kt{\widetilde{k}}
\def\dt{s}
\def\bI{\textbf{I}}
%% \def\dt{\widetilde{d}}
\def\Dt{\widetilde{D}}
\def\rk{\textrm{rank }}
\def\pa{\partial}
\def\D{\Delta}
\def\Z{\frak{Z}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\softO}{{O^{\sim}}}
\def\I{\frak{I}}
\def\Is{\frak{I}^{\star}}
\def\A{\frak{A}}
\def\fp{\frak{P}}
\def\sp{\mathscr{P}}
\def\la{\langle}
\def\ra{\rangle}
% matrices 
\def\scrQ{\ensuremath{\mathscr{Q}}}
\def\bbm{\begin{bmatrix}}
\def\ebm{\end{bmatrix}}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

Consider $f \in \ZZ[X_1,\hdots,X_n]$ with total degree $d$, and assume
that $f$ is squarefree and that $V(f) \subset \C^n$ is smooth. The key
to the proof of Theorem~\ref{theo:main} is the following quantitative
version of facts we stated above, namely that in generic coordinates,
polar varieties are smooth, equidimensional, and in Noether position.

We recall that an equidimensional algebraic set $X \subset \C^n$ of
dimension $d$ is in \textit{Noether position} for the projection
$\pi_d$ when the extension $\C[X_1,\hdots,X_d] \rightarrow
\C[X_1,\hdots,X_n]/I(X)$ is injective and integral; here, $I(X)
\subset \C[X_1,\dots,X_n]$ is the defining ideal of $X$. In this case,
for any $\xb \in \C^d,$ the fiber $X \cap \pi_d^{-1}(\xb)$ has
dimension zero (so it is finite and not empty).

For $i$ in $\{1,\dots,n-1\}$ and $f$ as above, we will let $\I(i, f)$
denote the sequence of $n-(i-1)$ polynomials $(f, {\pa f}/{\pa
  X_{i+1}},\dots,{\pa f}/{\pa X_n})$. As we pointed out in the
introduction, their zero-set is the $i$th polar variety $W(\pi_i,V(f))$.
Then, we say that $f$ satisfies $\textbf{H}_i$ if
\begin{enumerate}
\item For any $\xb$ in $W(\pi_i,V(f))$, the Jacobian matrix
  $\jac(\I(i, f))$ has full rank $n-(i-1)$ at $\xb$.

  {\em By the Jacobian Criterion~\cite[Corollary 16.20]{ECA}, this
  implies that $W(\pi_i,V(f))$ is either empty or $(i-1)$-equidimensional.}

@@

\item $W(\pi_i,V(f))$ is either empty or in Noether position for
  $\pi_{i-1}$.
\end{enumerate}
When this holds, given $\bm \alpha = (\alpha_1,\hdots,\alpha_{i-1})$
in $\C^{i-1}$, we further say that $f$ and $\bm \alpha$ satisfy
$\textbf{H}'_i$ if
\begin{enumerate}
\item the $n$ equations
  $(X_1-\alpha_1,\dots,X_{i-1}-\alpha_{i-1}),\I(i, f))$ generate a
  radical ideal.

@@

Even if $f$ does not satisfy $\textbf{H}_i$, we mentioned that it does
after applying a change of variables. The precise statement is the
following; to state it, we use the following notation. For a matrix
$\mA$ in $\C^{n^2}$ and $g$ in $\C[X_1,\hdots,X_n]$ we write
$g^\mA:=g(\mA \Xb) \in \C[X_1,\dots,X_n]$.

%% , $\xb \in
%% \mathbb{C}^n$ and $X \subset \C^n$. We write $\xb^\mA:=\mA^{-1}\xb$,
%% $X^\mA:=\{x^\mA~|~\xb \in V\}$ and $g^\mA:=g(\mA \Xb) \in
%% \C[X_1,\dots,X_n]$. In particular, we have $g^\mA(\xb^\mA) =g(\xb)$,
%% and $V(f)^\mA=V(f^\mA)$.

\begin{theorem}
  There exists a hypersurface $\D \subset \mathbb{C}^{n^2}$ of degree
  at most $\td$ such that, if $\mA$ is chosen from $\C^{n^2}-\Delta$
  then $\mA$ is invertible and $f^\mA$ satisfies $\textbf{H}_i$ for
  all $i \in \{1,\hdots,n-1\}.$
\end{theorem}
%% %
%% \begin{corollary} 
%%   Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1}
%%   D^{\textrm{\td}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen
%%   randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i
%%   \in \{1,\hdots,n\}$ is at least $1-\epsilon.$
%% \end{corollary}

It will on occasion be useful to work with change of variables with
``generic'' entries. For this, we will introduce $n^2$ new indeterminates
$(\frak A_{j,k})_{1\le j,k \le n^2}$, we will let $\K$ be the field
$\C(\frak A_{i,j})$, and use the corresponding
matrix $\A$ over $\K$. For $f$ as above, we will then define the
polynomial $f^\A:=f(\A \Xb) \in \K[X_1,\dots,X_n]$.

@@

Sard's lemma states that the set of critical values of a smooth
function $\R^n \to \R^m$ has measure zero. One can give ``algebraic''
versions of it, for semi-algebraic mappings $\R^n \to \R^m$ as
in~\cite[Chapter~9]{bochnak1998real}, or polynomial mappings $\C^n \to
\C^m$ as in~\cite[Chapter~3]{Mumford76}, for which the sets of
critical values are contained in strict semi-algebraic,
resp.\ algebraic sets in the codomain. Thom's weak transversality
lemma, as given for instance in~\cite{demazure2000bifurcations},
generalizes Sard's lemma. In this section, we consider a particular
case of this result (transversality to a point), and establish a
quantitative version of it; this will allow us to establish the first
item in property ${\bf H}_i$, as well as property ${\bf H}'_i$. Our
presentation is inspired by that in~\cite{TWT}.

%% ; it states that if a smooth function
%% $F(\Xb,\Lambda)$, depending on parameters $\Lamdba$ in $\C^s$ is
%% ``transverse'' to a certain submanifold in its domain space, then for
%% most values $\lambda \in \C^s$, $F(\Xb,\lambda)$ remains

@@

Transversality to a point can be rephrased entirely in terms of
critical and regular values. Recall that if $\bm \Psi$ is a mapping
from a smooth algebraic set $Y$ to $\C^t$, with $t\le \dim(Y)$, a {\em
  critical point} of $\bm \Psi$ is a point $\bm y \in Y$ such that the
image of the tangent space $T_{\bm y} Y$ by the differential $d_{\bm
  Y} \bm \Psi$ has dimension less than $t$. When for instance
$Y=\C^v$, we have $T_{\bm y} Y=\C^v$ and this condition is equivalent
to the Jacobian of $\bm \Psi$ having rank less than $t$ at $\bm y$.  A
{\em regular point} of $\bm \Psi$ is a point on $Y$ that is not
critical. {\em Critical values} are the images of critical points; the
complement of this set are the {\em regular values} (so a regular
value is not necessarily in the image of $\bm\Psi$).

Let $n,\dt,$ and $m$ be positive integers, with $m \le n$, and denote
by $\bm\Phi :\C^n \times \C^{\dt} ~ \rightarrow \C^{m}$ a mapping
defined by polynomials in $\C[\Xb,\Thetab]$, where $\Xb$,
resp.\ $\Thetab$, is a set of $n$, resp.\ $\dt$, indeterminates.  For
$\thetab$ in $\C^{\dt}$, let $\bm\Phi_{\thetab} : \C^n \rightarrow
\C^{m}$ be the induced mapping $\xb\mapsto \bm\Phi(\xb,\thetab)$.  The
transversality result we will need is the following.

\begin{proposition} [Weak transversality]\label{prop:weak_t}
  Suppose that $\bm 0$ is a regular value of $\bm\Phi$. Then there
  exists a hypersurface $\Delta \subset \C^{\dt}$ of degree at most
  \td for which, if $\thetab \in \C^{\dt}-\Delta$, then $\bm 0$ is a
  regular value of~$\bm\Phi_{\thetab}$.
\end{proposition}

Here is a simple example that shows this result at work. Consider a
squarefree $f$ in $\C[X_1,X_2]$, such that $V(f)$ is a smooth curve in
$\C^2$, and let the mapping $\bm\Phi:\C^2\times \C \to \C^2$ be defined
by $\bm\Phi(X_1,X_2,\Lambda) = (f(X_1,X_2), X_1-\Lambda)$. One checks
that the Jacobian of $\bm\Phi$ with respect to $(X_1,X_2,\Lambda)$ has
rank two at any point in $\bm\Phi^{-1}(\bm 0)$, so the assumptions of the
proposition apply. We deduce that for a generic $\lambda$ in $\C$,
that is, for all $\lambda$ in $\C$ except a finite number, the ideal
$(f(X_1,X_2), X_1-\lambda)$ is radical in $\C[X_1,X_2]$; equivalently,
$f(\lambda, X_2)$ is squarefree. We will revisit this example in
Section~\td.

The rest of the section is devoted to the proof of the proposition.
The proof of \cite[Theorem B.3]{TWT} already shows the existence of
$\Delta$; it is essentially the classical proof for smooth
mappings~\cite[Section~3.7]{demazure2000bifurcations}, written in an
algebraic context. In what follows, we revisit this proof,
establishing a bound on the degree of $\Delta$.

We will reuse the following fact proved in~\cite{TWT}.  Put $V =
\bm\Phi^{-1}(0)$ and consider the projection $\pi:(\xb, \thetab) \in
\C^n \times \C^{\dt} \mapsto \thetab \in\C^{\dt}$.  If $V$ is empty,
there is nothing to do, since all values $\thetab$ in $\C^\dt$ satisfy
the conclusion of the theorem. Thus, we assume that $V$ is not
empty. Then, the Jacobian criterion shows that $V$ is
smooth and $(n+ \dt -m)$-equidimensional.

Let $Z$ be the set of critical points of $\pi_{|V}$, and consider its
projection $\pi(Z)$ in $\C^s$. This is the set of critical values of
$\pi_{|V}$; hence, by the algebraic form of Sard's lemma (see
\cite[Theorem~3.7]{Mumford76} for irreducible $V$
and~\cite[Proposition~B.2]{TWT} for general $V$), its Zariski closure
$\overline{\pi(Z)}$ is a strict closed subset of $\C^s$. Now, the
proof of \cite[Theorem B.3]{TWT} shows that if $\thetab \in \C^{\dt}$
is not in $\overline{\pi(Z)}$, then $\bm 0$ is a critical value of
$\bm\Phi_{\thetab}$. Hence, it suffices to prove that $Z$ has degree
at most $d^{m+n}$; this will imply the same degree bound for
$\overline{\pi(Z)}$, and prove our propositon.

We first describe the set $Z$ of critical points of $\pi_{|V}$.  Let
$M$ denote the $(s+m) \times (s+n)$ matrix with entries in
$\C[\Xb,\Thetab]$ given by $M = \jac_{\Xb,\Thetab}(\pi,\bm\Phi)$, that
is,
\begin{align*}
M &= 
\bbm 
\jac_{\Xb,\Thetab}(\pi)\\
\jac_{\Xb,\Thetab}(\bm\Phi) 
\ebm 
=
\bbm 
\textbf{0}_{\dt \times n}\hspace{5mm}\textbf{I}_{\dt} \\
\jac_{(\Xb,\Thetab)}(\bm\Phi)
\ebm.
\end{align*}

\begin{lemma} 
  For $(\xb,\thetab)$ in $V$, the matrix $M$ has full rank $\dt+m$ at
  $(\xb,\thetab)$ if and only if $(\xb,\thetab)$ is a regular point of
  $\pi_{|V}.$
\end{lemma}
\begin{proof}
  Take $(\xb,\thetab)$ on $V$, and let $K(\xb,\thetab)$ be the
  Jacobian matrix $\jac_{(\Xb,\Thetab)}(\bm\Phi)$ at $(\xb,\thetab)$.
  Then, the rank of $M(\xb,\thetab)$ can be written as
  $\textup{rank}(K(\xb,\thetab)) + \textup{rank}([\textbf{0}_{\dt
      \times n}~\textbf{I}_{\dt}] \mid \ker K(\xb,\thetab))$, where
  the latter is the rank of the restriction of $[\textbf{0}_{\dt
      \times n}~\textbf{I}_{\dt}]$ to the nullspace of
  $K(\xb,\thetab)$.

  Since $V$ is smooth, $K(\xb,\thetab)$ has full rank ${\rm codim} (V)
  = m$. On the other hand, the nullspace of $K(\xb,\thetab)$ is the
  tangent space $T_{\xb,\thetab} V$, and
  $\textup{rank}([\textbf{0}_{\dt \times n}~\textbf{I}_{\dt}] \mid
  \ker K(\xb,\thetab))$ is the dimension of $\pi(T_{\xb,\thetab} V)$.
  In other words, the rank of $M(\xb,\thetab)$ is equal to
  $m+\dim(\pi(T_{\xb,\thetab} V))$; this implies the claim in the lemma.
\end{proof}

Therefore, we can characterize the set $Z$ of critical points of
$\pi_{|V}$ as those points satisfying $\bm\Phi(\xb,\thetab)=\bm 0$ and
where all minors of $M$ of order $\dt+m$ vanish. We can describe this
set using fewer equations, by discarding certain minors that are
identically zero. Let indeed $J$ denote the $m \times n$ submatrix of
the Jacobian of $\bm\Phi$ consisting of the first $n$ columns.
\begin{lemma} 
  For $(\xb,\thetab)$ in $V$, $(\xb,\thetab)$ is in $Z$ if and only if
  $J(\xb,\thetab)$ has rank less than $m$.
\end{lemma}
\begin{proof}
Notice 
%
\begin{align*}
M(\xb,\thetab)&= 
\bbm 
\textbf{0}_{\dt \times n} &\textbf{I}_{\dt} \\
J(\xb,\thetab)     &J'(\xb,\thetab)
\ebm,
\end{align*}
where $J'$ consists of the remaining columns of the Jacobian matrix of
$\bm\Phi$.  Then, the rank of the former matrix is equal to the rank
of
\begin{align*}
M(\xb,\thetab)&= 
\bbm 
\textbf{0}_{\dt \times n} &\textbf{I}_{\dt} \\
J(\xb,\thetab)     & \bm 0_{m \times s}
\ebm,
\end{align*}
and the conclusion follows.
\end{proof}
Therefore, the set $Z$ can be defined as the set of all
$(\xb,\thetab)$ in $\C^{n+\dt}$ satisfying $\bm\Phi(\xb,\thetab)=\bm
0$ and where all minors of $J$ of order $m$ vanish. Our next step is
to bound the degree of $Z$. For this, rather than considering minors
of $J$, we will rewrite the condition that $J(\xb,\thetab)$ has rank
less than $m$ as the existence of a non-trivial left kernel element.

For this, we let $\textit{\textbf{L}}=[L_1,\hdots,L_m]$ be new
variables, thought of as Lagrange multipliers, and consider the
``Lagrange polynomials'' $\mathscr{L}_1,\dots,\mathscr{L}_{n}$, with
\[
[\mathscr{L}_1 \cdots \mathscr{L}_{n}]=\textit{\textbf{L}}\cdot J(\xb,\thetab).
\] 
Denote by $\frak{Z} \subset \C^{n+\dt+m}$ the algebraic set defined by
the vanishing of $\mathscr{L}_1,\hdots,\mathscr{L}_{n},$ and
$\bm\Psi$, and by $\frak{Z}'$ the algebraic set
\[
\frak{Z}' := \overline{\frak{Z} - \{(\xb,\thetab,0,\hdots,0) \in \C^{n+\dt+m}~|~(\xb,\thetab,0\hdots,0) \in \frak{Z}\}},
\]
where the bar denotes Zariski closure (we have to remove such points,
since $L_1=\cdots=L_m=0$ is always a trivial solution to the Lagrange
equations). Finally, consider the projection
\begin{align*} 
\mu :~ \C^{n+\dt+m} &\rightarrow \C^{n+\dt}\\
(\xb,\thetab,\bm \ell)~ &\mapsto (\xb,\thetab).
\end{align*}

>>>>

% packages 
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[titles]{tocloft}%----------------------------------------
\usepackage{mathrsfs}
\usepackage{bm}
\usepackage{enumitem}
% new commands and shortcuts
\def\td{{\bf todo}}
\def\bs{\textit{\textbf{s}}}
\def\bz{\textit{\textbf{0}}}
\def\Lb{\textit{\textbf{L}}}
\def\Xb{\textit{\textbf{X}}}
\def\Lambdab{\bm{\Lambda}}
\def\Thetab{\bm{\Theta}}
\def\thetab{\bm{\vartheta}}
\def\mA{{\bm A}}
\def\fA{{\frak A}}
\def\xb{\textit{\textbf{x}}}
\def\fb{\textit{\textbf{f}}}
\def\ab{\textit{\textbf{a}}}
\def\pb{\textit{\textbf{p}}}
\def\ajb{\overline{a_j}}
\def\bjb{\overline{b_{j,l}}}
\def\pjb{\overline{P_{j}}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\htt}{ht}
\def\C{\mathbb{C}}
\def\pr{\mathbb{P}}
\def\vt{\vartheta}
\def\Q{\mathbb{Q}}
\def\R{\mathbb{R}}
\def\K{\mathbb{K}}
\def\rs{\mathscr{R}}
\def\P{\mathscr{P}}
\def\d{\delta}
\def\sing{ \textrm{sing}}
\def\codim{ \textrm{codim}}
\def\jac{ \textbf{jac}}
\def\exp{\textrm{exp}}
\def\grad{\textrm{grad}}
\def\rank{\textrm{rank}}
\def\reg{\textrm{reg}}
\def\jt{\widetilde{J}}
\def\At{\widetilde{A}}
\def\Yt{\widetilde{Y}}
\def\dtt{\widetilde{d}}
\def\kt{\widetilde{k}}
\def\dt{s}
\def\bI{\textbf{I}}
%% \def\dt{\widetilde{d}}
\def\Dt{\widetilde{D}}
\def\rk{\textrm{rank }}
\def\pa{\partial}
\def\D{\Delta}
\def\Z{\frak{Z}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\softO}{{O^{\sim}}}
\def\I{\frak{I}}
\def\Is{\frak{I}^{\star}}
\def\A{\frak{A}}
\def\fp{\frak{P}}
\def\sp{\mathscr{P}}
\def\la{\langle}
\def\ra{\rangle}
% matrices 
\def\scrQ{\ensuremath{\mathscr{Q}}}
\def\bbm{\begin{bmatrix}}
\def\ebm{\end{bmatrix}}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

Consider $f \in \ZZ[X_1,\hdots,X_n]$ with total degree $d$, and assume
that $f$ is squarefree and that $V(f) \subset \C^n$ is smooth. The key
to the proof of Theorem~\ref{theo:main} is the following quantitative
version of facts we stated above, namely that in generic coordinates,
polar varieties are smooth, equidimensional, and in Noether position
(or empty).

We recall that an equidimensional algebraic set $X \subset \C^n$ of
dimension $d$ is in \textit{Noether position} for the projection
$\pi_d$ when the extension $\C[X_1,\hdots,X_d] \rightarrow
\C[X_1,\hdots,X_n]/I(X)$ is integral; here, $I(X) \subset
\C[X_1,\dots,X_n]$ is the defining ideal of $X$. In this case, for any
$\xb \in \C^d,$ the fiber $X \cap \pi_d^{-1}(\xb)$ has dimension zero
(so it is finite and not empty).

For $i$ in $\{1,\dots,n\}$ and $f$ as above, we will let $\I(i, f)$
denote the sequence of $n-(i-1)$ polynomials $(f, {\pa f}/{\pa
  X_{i+1}},\dots,{\pa f}/{\pa X_n})$. As pointed out in the
introduction, their zero-set is the $i$th polar variety
$W(\pi_i,V(f))$.  Then, we say that $f$ satisfies $\textbf{H}_i$ if
\begin{enumerate}
\item For any $\xb$ in $W(\pi_i,V(f))$, the Jacobian matrix
  $\jac(\I(i, f))$ has full rank $n-(i-1)$ at $\xb$.

  {\em By the Jacobian Criterion~\cite[Corollary 16.20]{ECA}, this
  implies that $W(\pi_i,V(f))$ is either empty or $(i-1)$-equidimensional,
  and that $\I(i, f^\mA)$ is radical.}

@@

\item $W(\pi_i,V(f))$ is either empty or in Noether position for
  $\pi_{i-1}$.
\end{enumerate}
When this holds, given $\bm \alpha = (\alpha_1,\hdots,\alpha_{i-1})$
in $\C^{i-1}$, we further say that $f$ and $\bm \alpha$ satisfy
$\textbf{H}'_i$ if
\begin{enumerate}
\item The $n$ polynomials
  $(X_1-\alpha_1,\dots,X_{i-1}-\alpha_{i-1},\I(i, f))$ generate a
  radical ideal.

@@

Even if $f$ does not satisfy $\textbf{H}_i$, we mentioned that it
does, after applying a generic change of variables. The precise
statement is the following; to state it, we use the following
notation. For a matrix $\mA$ in $\C^{n\times n}$ and $g$ in
$\C[X_1,\hdots,X_n]$ we write $g^\mA:=g(\mA \Xb) \in
\C[X_1,\dots,X_n]$, where $\Xb$ is the column vector with entries
$X_1,\dots,X_n$.

We will also have to consider matrices with generic entries. For this,
we introduce $n^2$ new indeterminates $(\frak A_{j,k})_{1\le j,k \le
  n}$. Then, $\A$ will denote the matrix with entries $(\frak
A_{j,k})_{1\le j,k \le n}$, $\C(\A)$ will denote the rational function
field $\C((\frak A_{j,k})_{1\le j,k \le n})$ and $\C[\A]$ the
polynomial ring $\C[(\frak A_{j,k})_{1\le j,k \le n}]$.  For $f$ as
above, we will then define the polynomial $f^\A:=f(\A \Xb)$, 
which we will consider in either
$\C(\A)[X_1,\dots,X_n]$ or $\C[\A,X_1,\dots,X_n]$.

\begin{theorem}\label{theo:gen}
  For $i=1,\dots,n$, there exists a non-zero polynomial $\D_i$ in
  $\C[\A]$ of degree at most $\td$ such that if $\mA \in \C^{n\times
    n}$ does not cancel $\D_i$, then $\mA$ is invertible and $f^\mA$
  satisfies $\textbf{H}_i$.
\end{theorem}
\noindent The proof of this theorem occupies the next two sections.

@@

Sard's lemma states that the set of critical values of a smooth
function $\R^n \to \R^m$ has measure zero. One can give ``algebraic''
versions of it, for semi-algebraic mappings $\R^n \to \R^m$ as
in~\cite[Chapter~9]{bochnak1998real}, or polynomial mappings $\C^n \to
\C^m$ as in~\cite[Chapter~3]{Mumford76}, for which the sets of
critical values are contained in strict semi-algebraic,
resp.\ algebraic sets in the codomain. Thom's weak transversality
lemma, as given for instance in~\cite{demazure2000bifurcations},
generalizes Sard's lemma. In this section, we consider a particular
case of this result (transversality to a point), and establish a
quantitative version of it; this will allow us to establish the first
item in property ${\bf H}_i$, as well as property ${\bf H}'_i$.

@@

Transversality to a point can be rephrased entirely in terms of
critical and regular values. Recall that if $\bm \Psi$ is a mapping
from a smooth algebraic set $Y$ to $\C^t$, with $t\le \dim(Y)$, a {\em
  critical point} of $\bm \Psi$ is a point $\bm y \in Y$ such that the
image of the tangent space $T_{\bm y} Y$ by the differential $d_{\bm
  Y} \bm \Psi$ has dimension less than $t$. When for instance
$Y=\C^v$, we have $T_{\bm y} Y=\C^v$ and this condition is equivalent
to the Jacobian of $\bm \Psi$ having rank less than $t$ at $\bm y$.
{\em Critical values} are the images by $\Psi$ of critical points; the
complement of this set are the {\em regular values} (so a regular
value is not necessarily in the image of~$\bm\Psi$).

Let then $n,\dt,$ and $m$ be positive integers, with $m \le n$, and denote
by $\bm\Phi :\C^n \times \C^{\dt} ~ \rightarrow \C^{m}$ a mapping
defined by polynomials in $\C[\Xb,\Thetab]$, where $\Xb$,
resp.\ $\Thetab$, is a set of $n$, resp.\ $\dt$, indeterminates.  For
$\thetab$ in $\C^{\dt}$, let $\bm\Phi_{\thetab} : \C^n \rightarrow
\C^{m}$ be the induced mapping $\xb\mapsto \bm\Phi(\xb,\thetab)$.  The
transversality result we will need is the following.

\begin{proposition} [Weak transversality]\label{prop:weak_t}
  Suppose that $\bm 0$ is a regular value of $\bm\Phi$. Then there
  exists a non-zero polynomial $\Gamma \in \C[\Thetab]$ of degree at
  most $d^{m+n}$ such that for $\thetab$ in $\C^\dt$, if
  $\Gamma(\thetab)\ne 0$, then $\bm 0$ is a regular value
  of~$\bm\Phi_{\thetab}$.
\end{proposition}

Here is a simple example that shows this result at work. Consider a
squarefree $f$ in $\C[X_1,X_2]$, such that $V(f)$ is a smooth curve in
$\C^2$, and let the mapping $\bm\Phi:\C^2\times \C \to \C^2$ be
defined by $\bm\Phi(X_1,X_2,\Theta) = (f(X_1,X_2), X_1-\Theta)$. One
checks that the Jacobian of $\bm\Phi$ with respect to
$(X_1,X_2,\Theta)$ has rank two at any point in $\bm\Phi^{-1}(\bm 0)$,
so the assumptions of the proposition apply. We deduce that for a
generic $\vartheta$ in $\C$, that is, for all $\vartheta$ in $\C$
except a finite number, the ideal $(f(X_1,X_2), X_1-\vartheta)$ is
radical in $\C[X_1,X_2]$; equivalently, $f(\vartheta, X_2)$ is
squarefree. We will revisit this example in Section~\ref{ssec:fiber}.

The rest of the subsection is devoted to the proof of the proposition.
The proof of \cite[Theorem B.3]{TWT} already shows the existence of
$\Gamma$; it is essentially the classical proof for smooth
mappings~\cite[Section~3.7]{demazure2000bifurcations}, written in an
algebraic context. In what follows, we revisit this proof,
establishing a bound on the degree of $\Gamma$.

Put $V = \bm\Phi^{-1}(\bm 0)$. If $V$ is empty, there is nothing to do,
since all values $\thetab$ in $\C^\dt$ satisfy the conclusion of the
proposition. Thus, we assume that $V$ is not empty. Then, the Jacobian
criterion shows that $V$ is smooth and $(n+ \dt -m)$-equidimensional.

We will reuse the following fact, proved in~\cite{TWT}.  Consider the
projection $\pi:(\xb, \thetab) \in \C^n \times \C^{\dt} \mapsto
\thetab \in\C^{\dt}$.  Let $Z$ be the set of critical points of
$\pi_{|V}$, and consider its projection $\pi(Z)$ in $\C^s$. This is
the set of critical values of $\pi_{|V}$; hence, by the algebraic form
of Sard's lemma (see \cite[Theorem~3.7]{Mumford76} for irreducible $V$
and~\cite[Proposition~B.2]{TWT} for general $V$), its Zariski closure
$\overline{\pi(Z)}$ is a strict closed subset of $\C^s$. As we will
see below, if $\thetab \in \C^{\dt}$ is not in $\overline{\pi(Z)}$,
then $\bm 0$ is a regular value of $\bm\Phi_{\thetab}$.

% Now, the proof of \cite[Theorem B.3]{TWT} shows that if $\thetab \in
% \C^{\dt}$ is not in $\overline{\pi(Z)}$, then $\bm 0$ is a regular
% value of $\bm\Phi_{\thetab}$.

To describe the set $Z$ of critical points of $\pi_{|V}$, let $\bm M$
denote the $(s+m) \times (s+n)$ Jacobian matrix with entries in
$\C[\Xb,\Thetab]$ given by $\bm M = \jac_{\Xb,\Thetab}(\pi,\bm\Phi)$,
that is,
\begin{align*}
\bm M &= 
\bbm 
\jac_{\Xb,\Thetab}(\pi)\\
\jac_{\Xb,\Thetab}(\bm\Phi) 
\ebm 
=
\bbm 
\textbf{0}_{\dt \times n}\hspace{5mm}\textbf{I}_{\dt} \\
\jac_{\Xb,\Thetab}(\bm\Phi)
\ebm.
\end{align*}

\begin{lemma} 
  For $(\xb,\thetab)$ in $V$, $(\xb,\thetab)$ is in $Z$ if and only if
  the matrix $\bm M$ has rank less than $\dt+m$ at $(\xb,\thetab)$.
\end{lemma}
\begin{proof}
  Take $(\xb,\thetab)$ on $V$, and let $\bm K(\xb,\thetab)$ be the
  Jacobian matrix $\jac_{\Xb,\Thetab}(\bm\Phi)$ taken at $(\xb,\thetab)$.
  Then, the rank of $\bm M(\xb,\thetab)$ can be written as
  $\textup{rank}(\bm K(\xb,\thetab)) + \textup{rank}([\textbf{0}_{\dt
      \times n}~\textbf{I}_{\dt}] \mid \ker \bm K(\xb,\thetab))$, where
  the latter is the rank of the restriction of $[\textbf{0}_{\dt
      \times n}~\textbf{I}_{\dt}]$ to the nullspace of
  $\bm K(\xb,\thetab)$.

  Since $V$ is smooth, $\bm K(\xb,\thetab)$ has full rank ${\rm codim}
  (V) = m$. On the other hand, the nullspace of $\bm K(\xb,\thetab)$
  is the tangent space $T_{\xb,\thetab} V$, and
  $\textup{rank}([\textbf{0}_{\dt \times n}~\textbf{I}_{\dt}] \mid
  \ker \bm K(\xb,\thetab))$ is the dimension of $\pi(T_{\xb,\thetab}
  V)$.  In other words, the rank of $\bm M(\xb,\thetab)$ is equal to
  $m+\dim(\pi(T_{\xb,\thetab} V))$; this implies the claim in the
  lemma.
\end{proof}

Therefore, we can characterize the set $Z$ of critical points of
$\pi_{|V}$ as those points satisfying $\bm\Phi(\xb,\thetab)=\bm 0$ and
where all minors of $\bm M$ of order $\dt+m$ vanish. We can actually
describe this set using a smaller matrix, by discarding certain minors
that are identically zero. Let indeed $\bm J$ denote the $m \times n$
submatrix of the Jacobian of $\bm\Phi$ consisting of the first $n$
columns. This is the Jacobian matrix of $\bm \Phi$ with respect to $\Xb$.
\begin{lemma} 
  For $(\xb,\thetab)$ in $V$, $(\xb,\thetab)$ is in $Z$ if and only if
  $\bm J(\xb,\thetab)$ has rank less than $m$.
\end{lemma}
\begin{proof}
Notice 
%
\begin{align*}
M(\xb,\thetab)&= 
\bbm 
\textbf{0}_{\dt \times n} &\textbf{I}_{\dt} \\
\bm J(\xb,\thetab)     &\bm J'(\xb,\thetab)
\ebm,
\end{align*}
where $\bm J'$ consists of the remaining columns of the Jacobian matrix of
$\bm\Phi$.  Then, the rank of the former matrix is equal to the rank
of
\begin{align*}
M(\xb,\thetab)&= 
\bbm 
\textbf{0}_{\dt \times n} &\textbf{I}_{\dt} \\
\bm J(\xb,\thetab)     & \bm 0_{m \times s}
\ebm,
\end{align*}
and the conclusion follows.
\end{proof}

In particular, take $\thetab$ in $\C^\dt - \overline{\pi(Z)}$.
Then for all $\xb$ in $\bm \Phi_{\thetab}^{-1}(\bm 0)$, $(\xb,\thetab)$
is in $V$, so it is not in $Z$. The previous lemma then implies that
the Jacobian matrix $\bm J$ of $\bm\Phi_{\thetab}$ has full rank $m$ at
$(\xb,\thetab)$. In other words, $\bm 0$ is a regular value of $\bm
\Phi_{\thetab}$, as claimed.

Our next step is to bound the degree of $Z$.  In that, we use the
definition of degree given in~\cite{H}: the degree of an irreducible
algebraic set is the number of intersection points it has with a
generic hyperplane of complementary dimension, and the degree of an
arbitrary algebraic set is the sum of the degrees of its irreducible
components.  To obtain an estimate on the degree of $Z$, rather than
considering minors of $\bm J$, we will rewrite the condition that $\bm
J(\xb,\thetab)$ has rank less than $m$ as the existence of a
non-trivial left kernel element.
